---
title: "Modeling Financial Ranges with Intervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling Financial Ranges with Intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(boostedintervals)
# finance examples use dimensionless rates (unit = "1")
```

## Motivation

Financial inputs like interest rates or returns are often uncertain. Intervals
can express ranges of plausible rates and propagate them through compounding.

## Compound growth example

Principal grows by an uncertain annual return `r` for `n` years:

\[ FV = P (1 + r)^n \]

```{r}
P <- units_interval(1000, 1000, unit = "1")  # unitless money placeholder
r <- units_interval(0.03, 0.05, unit = "1")   # 3% to 5%
n <- 5

fv <- P * (1 + r)^n
as_numeric_bounds(fv)
```

## Decision thresholds

We can check whether a target outcome is possible or certain.

```{r}
target <- units_interval(1200, 1200, unit = "1")
possible(fv, target, ">=")
certain(fv, target, ">=")
```

## Takeaways

- Treat rates as dimensionless intervals.
- Convert to numeric bounds for reporting; consider currency formatting outside
  the package.

---

