---
title: "Spatial and Measurement Uncertainty in Geoscience"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial and Measurement Uncertainty in Geoscience}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(boostedintervals)
library(units)
```

## Motivation

Sensors and surveys produce measurements with uncertainty. Representing
elevation, depth, and distance as intervals makes downstream spatial analysis
aware of measurement error.

## Elevation example

Two GPS-derived elevation measurements with uncertainty.

```{r}
site_a <- units_interval(set_units(120.0, "m"), set_units(120.5, "m"))
site_b <- units_interval(set_units(150.2, "m"), set_units(150.8, "m"))

# Distance in vertical (simple example)
dist <- abs(site_b - site_a)
midpoint(dist)
as_numeric_bounds(dist, unit = "m")
```

## Inflating measurement uncertainty for conservative planning

We may wish to inflate uncertainties to account for unknown biases.

```{r}
inflated <- inflate(site_a, relative = 0.1)
inflated
```

## Takeaways

- Keep units attached to prevent accidental mixing of incompatible measures.
- Interval diagnostics support conservative planning and risk-aware mapping.

---

