---
title: "Propagating Uncertainty in Physics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Propagating Uncertainty in Physics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(boostedintervals)
library(units)
```

## Motivation

Small measurement uncertainties in mass and velocity propagate nonlinearly
into derived physical quantities such as kinetic energy. Intervals allow us to
represent and propagate these uncertainties while keeping unit metadata.

## Kinetic energy example

The kinetic energy of a mass `m` moving at speed `v` is:

\[ E = \tfrac{1}{2} m v^2 \]

We build unit-aware intervals for `m` and `v`, then compute `E` and extract
numeric bounds.

```{r}
mass <- units_interval(set_units(1.0, "kg"), set_units(1.02, "kg"))
vel  <- units_interval(set_units(10.0, "m/s"), set_units(10.2, "m/s"))
energy <- 0.5 * mass * vel^2
energy
as_numeric_bounds(energy, unit = "J")
```

## Diagnostics and safety check

We can ask whether energy may exceed a safety threshold or whether it is
certainly below it.

```{r}
threshold <- units_interval(set_units(50, "J"), set_units(55, "J"))
possible(energy, threshold, ">=")
certain(energy, threshold, ">=")
verify(energy, threshold, ">=")
```

## Takeaways

- Intervals propagate non-linear uncertainty naturally.
- Units remain attached, and conversion to numeric bounds is straightforward
  for downstream reporting.

---

